Applying Pose Estimation to Predict the Outcome of Basketball Shots

### 2022-11-28 Papers/Report/report.tex
Add general overview as well as data collection section

Pose estimation attempts to detect the location of 17 keypoints on a human body, as seen in \ref{fig:pose}. These keypoints include key joints such as
the knee, elbow, shoulder, and wrist, which can be vital in determining the form of a free throw.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.25\textwidth]{imgs/openpose.png}
  \caption{Pose estimation keypoints}
  \label{fig:pose}
\end{figure}
%

@@

%------------------------------------------------------------------------
\section{Methodology}
\subsection{General Overview}
We will have a camera taking in video input of a player shooting the basketball, potentially from two different angles.
One angle would be a side view of the shooter that can capture the relative positions of the shooter and the hoop, and the other
would be a head-on view of the of the shooter to capture the full pose estimation of the shooter.

In the background, we will have our pose estimation algorithm running to capture the pose estimation of the shooter which will be used
to make our predictions. Additionally, we will potentially be retrieving other data to create our feature vector, including the angle relationship
between the shooter and the hoop, and the speed of the shot. We will be attempting to capture the pose estimation at multiple stages of the
shot, as each stage can have a significant impact on the outcome of the shot. We will then take this feature vector and put it through a classification model.
We will perform binary classification and predict whether a shot will go in or not.

>>>>

Pose estimation attempts to detect the location of 17 keypoints on a human body, as seen in Figure \ref{fig:pose}. These keypoints include key joints such as
the knee, elbow, shoulder, and wrist, which can be vital in determining the form of a free throw.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.25\textwidth]{imgs/openpose.png}
  \caption{Pose estimation keypoints}
  \label{fig:pose}
\end{figure}
%

@@

%------------------------------------------------------------------------
\section{Methodology}
\subsection{General Overview}
% We will have a camera taking in video input of a player shooting the basketball, potentially from two different angles.
% One angle would be a side view of the shooter that can capture the relative positions of the shooter and the hoop, and the other
% would be a head-on view of the of the shooter to capture the full pose estimation of the shooter.

% In the background, we will have our pose estimation algorithm running to capture the pose estimation of the shooter which will be used
% to make our predictions. Additionally, we will potentially be retrieving other data to create our feature vector, including the angle relationship
% between the shooter and the hoop, and the speed of the shot. We will be attempting to capture the pose estimation at multiple stages of the
% shot, as each stage can have a significant impact on the outcome of the shot. We will then take this feature vector and put it through a classification model.
% We will perform binary classification and predict whether a shot will go in or not.
The high-level overview of our end-to-end pipeline is shown in Figure \ref{fig:flowchart}. We start off by capturing clips of a test subject shooting a free throw. This will be our 
original data. We first want to downsample the video as we assume that many of the frames will be redundant and take up unnecessary computation time. After this, we perform pose estimation
on all the frames of each video to get the keypoints of our test subject for each video clip. Using the pose estimation data, we perform
feature extraction to generate feature vectors for each video clip according to some convention. This will be elaborated on in a later section.

These feature vectors will be the training and testing data that we use for our classification model. We will want to split this data into training
and testing sets. We will train our classification model on the training set and generate predictions using our testing set.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{imgs/cv_pipeline_real.png}
  \caption{Overview of Pipeline}
  \label{fig:flowchart}
\end{figure}

\subsection{Data Collection}
To fit the scope of our project, we wanted the shot data be as controlled as possible to prevent unwanted variations
in our data. To do this, we collected our own data, where we had our test subject always shoot from the free throw line
and had a camera in a fixed position in front of the test subject to capture their form. The set up can be seen in Figure \ref{fig:ft_data}.
We genereated approxiamately 300 clips of our test subject shooting free throws. Additionally, for this project, we wanted to detect major changes
in shooting forms that could affect the outcome of a shot. Therefore, our test subject shot some of their shots as he normally would, with the intent
of making it, and some of their shots with purposefully bad form, with the intent of missing. We had a relatively even split of these two types of shots.

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{imgs/data_collection.png}
  \caption{Free Throw}
  \label{fig:ft_data}
\end{figure}
